%&thesis
\documentclass[a4paper,11pt,BCOR=8mm,twoside,headsepline]{scrbook}
\author{Daniel Suess}
\title{Due to, or in spite of? The effect of constraints on efficiency in quantum estimation problems}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Begin of dumped preamble
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \usepackage[T1]{fontenc}
\usepackage{ifpdf}
\ifpdf{}
  \usepackage[utf8]{inputenc}
\fi
\usepackage{ifluatex}
\ifluatex{}
  \usepackage[utf8]{luainputenc}
  \usepackage{shellesc}
\fi

\usepackage{graphicx,subcaption}
\usepackage[backend=biber,style=alphabetic,bibencoding=utf8]{biblatex}
\bibliography{references_ds}

\usepackage[color=red!50!white,textsize=small,textwidth=2.5cm]{todonotes}

% Externalize tikz pictures except for todonotes
\usepackage{tikz}%for the figures
\usetikzlibrary{external,calc,backgrounds}
\tikzexternalize[prefix=build/]
\tikzset{external/mode=graphics if exists}
% \tikzset{external/system call={lualatex \tikzexternalcheckshellescape -halt-on-error -interaction=batchmode -jobname "\image" "\texsource"}}


\makeatletter
\renewcommand{\todo}[2][]{\tikzexternaldisable\@todo[noline,#1]{#2}\tikzexternalenable}
\makeatother
\usepackage{enumitem}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[breaklinks=true]{hyperref}
%getting rid of hyperref's ugly boxes.
%From:http://tex.stackexchange.com/a/51349
\hypersetup{
  colorlinks   = true, %Colours links instead of ugly boxes
  urlcolor     = blue, %Colour for external hyperlinks
  linkcolor    = blue, %Colour of internal links
  citecolor   = red %Colour of citations
}

% if \makefull is defined (e.g. by commandline argument, include everything)
\ifdefined\makefull\else
  \includeonly{chapters/introduction,chapters/conclusion}
  % \includeonly{chapters/error_region,chapters/error_region_appendix}
  % \includeonly{chapters/phaselift,chapters/phaselift_appendix}
  % \includeonly{chapters/tensors,chapters/tensors_appendix}
\fi

\input{macros}
\input{notebook-macros}

%% use \[ ... \] for displayed equations
\newcommand{\myequation}{\begin{equation}}
\newcommand{\myendequation}{\end{equation}}
\let\[\myequation{}
\let\]\myendequation{}

%% Only referenced equations are numbered
% \usepackage{autonum}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\frontmatter
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter*{Abstract}

In this thesis, we study the interplay of constraints and complexity in quantum estimation.
We investigate three inference problems, where additional structure in the form of constraints is exploited to reduce the sample and/or computational complexity.
The first example is concerned with uncertainty quantification in quantum state estimation, where the positive-semidefinite constraint is used to construct more powerful, that is smaller, error regions.
However, as we show in this work, doing so in an optimal way constitutes a computationally hard problem, and therefore, is intractable for larger systems.
This is in stark contrast to the unconstrained version of the problem under consideration.
The second inference problem deals with phase retrieval in general, and characterizing linear optical circuits in particular.
The main challenge here is the fact that the measurements are oblivious to phases, and hence, their reconstruction requires deliberate interference.
We propose a reconstruction algorithm based on ideas from low-rank matrix recovery, which exploits an exact rank-one constraint on the signal to be  recovered.
Furthermore, in this work we propose a measurement ensemble tailored to the specific application of characterizing linear optical devices.
Finally, we investigate low-rank tensor recovery -- the problem of reconstructing a low-complexity tensor embedded in an exponentially large space.
We provide evidence that a low-rank tensor can be recovered from a number of measurements that only depends on its intrinsic complexity, and hence, scales polynomially in the order of tensor.
Therefore, the low-rank constraint can be exploited to dramatically reduce the sample complexity of the problem.
By using efficiently representable measurement tensors, our approach is also computationally efficient.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter*{Kurzzusammenfassung}

\todo[inline]{TODO}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\tableofcontents
\newpage
\makeatletter
\providecommand\@dotsep{5}
\makeatother
\listoftodos\relax
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\mainmatter

\include{chapters/introduction}
\include{chapters/error_region}
\include{chapters/phaselift}
\include{chapters/tensors}
\include{chapters/conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix

\chapter{Appendix}%
\include{chapters/error_region_appendix}
\include{chapters/phaselift_appendix}
\include{chapters/tensors_appendix}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\printbibliography[heading=bibintoc]


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter*{Acknowledgments}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter*{Erklärung}

Ich versichere, dass ich die von mir vorgelegte Dissertation selbständig angefertigt, die benutzten Quellen und Hilfsmittel vollständig angegeben und die Stellen der Arbeit -- einschließlich Tabellen, Karten und Abbildungen --, die anderen Werken im Wortlaut oder dem Sinn nach entnommen sind, in jedem Einzelfall als Entlehnung kenntlich gemacht habe; dass diese Dissertation noch keiner anderen Fakultät oder Universität zur Prüfung vorgelegen hat; dass sie -- abgesehen von unten angegebenen Teilpublikationen -- noch nicht veröffentlicht worden ist, sowie, dass ich eine solche Veröffentlichung vor Abschluss des Promotionsverfahrens nicht vornehmen werde.
Die Bestimmungen der Promotionsordnung sind mir bekannt.
Die von mir vorgelegte Dissertation ist von Prof.\ Dr.\ David Gross betreut worden.

\vspace{4cm}

\hspace{2cm} Ort, Datum \hfill Unterschrift \hspace{2cm}

\vspace{\fill}
\subsection*{Teilpublikationen}
\begin{itemize}
  \item D.\ Suess, Ł.\ Rudnicki, T.\ O.\ Maciel, D.\ Gross: \textit{Error regions in quantum state tomography: computational complexity caused by geometry of quantum states}, New J.\ Phys.\ 19 093013 (2017)
  \item Ž.\ Stojanac, D.\ Suess, M.\ Kliesch: \textit{On the distribution of a product of N Gaussian random variables}, Proceedings Volume 10394, Wavelets and Sparsity XVII; 1039419 (2017)
  \item Ž.\ Stojanac, D.\ Suess, M.\ Kliesch, \textit{On products of Gaussian random variables}, arXiv:1711.10516
  \item D.\ Suess, M.\ Holzaepfel, \textit{mpnum: A matrix product representation library for Python}, Journal of Open Source Software, 2(20), 465 (2017)
\end{itemize}


\end{document}
